When Do You Repeat Yourself?\\ Voices from the Trenches of Linux Kernel Maintainers on Code Duplication

### 2025-10-28 papers/techdebt2026/techdebt.tex
techdebt2026: techdebt: standardize the term "coding style"
Throughout the text, there are two instances of "code style", which
isn't incorrect, yet, it diverges from what we were using ("coding
style"). In this sense, standardize this term, by replacing the former
occurences with the latter.

Signed-off-by: David Tadokoro <davidbtadokoro@usp.br>

The 2025 course offering had 37 students (25 undergraduate and 12 graduate), who were asked to form groups of two or three members, but graduate students could work individually. To assist students in this task, we (a professor and three more experienced graduate students) prepared several alternatives for contributing to the Linux kernel, including simpler contributions such as fixing code style issues. In this context, we specified two options for removing duplications in the Linux kernel.

@@

In this attempt, we reached a point where the deduplication was good enough to submit it as a contribution to the AMD Display driver code quality. Thus, we moved to sending the refactoring to the driver maintainers as a patch.

%TODO: dado que falamos da demora do feedback, colocar a data de envia da V1.
In the feedback on the first submitted version, the maintainers only asked for minor changes to align with the coding style, correct license use, and align with initially unknown conventions. We submitted a second version, in which the maintainers asked us to move the functions created in the generic library to an existing file instead of creating a new one. We sent a third version, and the patch was accepted and integrated into the kernel codebase on February 25, 2025.
%\footnote{Patch submitted: \href{https://lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br/}{lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br}}.

@@

Groups 7 and 8 used the extract method to approach the deduplications. Group 8 sent their patch to the maintainers, and the code changes were approved, with minor requested changes in code style and smaller errors fixed (++T5). In the case of Group 7, the maintainer initially rejected the proposed patch but suggested a new refactoring approach: applying the extraction method refactoring to the duplicated code to an existing helper function. After the feedback cycle, maintainers accepted the patch in version 3.

>>>>

The 2025 course offering had 37 students (25 undergraduate and 12 graduate), who were asked to form groups of two or three members, but graduate students could work individually. To assist students in this task, we (a professor and three more experienced graduate students) prepared several alternatives for contributing to the Linux kernel, including simpler contributions such as fixing coding style issues. In this context, we specified two options for removing duplications in the Linux kernel.

@@

At this stage, we considered the deduplication proposal good enough to submit it to the AMD Display driver as a patch. We sent the first patch on August 9, 2025, and a resend followed on October 9, 2025, due to a lack of initial feedback. Nevertheless, for the first two versions, the maintainers' response requested minor changes for coding style, license use, and alignment with existing conventions. In the third version, the feedback was to move the new generic library functions to an existing file, instead of creating a new one. Then, the fourth version was accepted and integrated into the kernel on February 25, 2025.
%\footnote{V1: \href{https://lore.kernel.org/amd-gfx/20240809210122.23814-1-luanicaro@usp.br/}{https://lore.kernel.org/amd-gfx/20240809210122.23814-1-luanicaro@usp.br/}, V1 (resend): \href{https://lore.kernel.org/amd-gfx/20241009153424.15256-1-luanicaro@usp.br/}{https://lore.kernel.org/amd-gfx/20241009153424.15256-1-luanicaro@usp.br/}, V4 (approved): \href{https://lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br/}{lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br/}}.

@@

Groups 7 and 8 used the extract method to approach the deduplications. Group 8 sent their patch to the maintainers, and the code changes were approved, with minor requested changes in coding style and smaller errors fixed (++T5). In the case of Group 7, the maintainer initially rejected the proposed patch but suggested a new refactoring approach: applying the extraction method refactoring to the duplicated code to an existing helper function. After the feedback cycle, maintainers accepted the patch in version 3.
